---
title: "Data retrival and wrangling"
author: "Irina Kalatskaya"
date: "`r Sys.Date()`"
always_allow_html: true
output: 
  md_document:
    toc: true
   
---

```{r setup, include=FALSE}

library(rvest)
library(stringr)
library(modelr)
library(dplyr, warn.conflicts = FALSE)
library(readr)
library(tidyverse)

```

<br/>

## House rent price in MA

The data was scraped from https://www.rentdata.org/states/massachusetts/.

```{r rent, warning=FALSE}

rentMA = tibble()
years = seq(from = 2006, to = 2022)

for(year in years) {
  
  url = paste0('https://www.rentdata.org/states/massachusetts/', year)
  print(url)
  #Reading the HTML code from the website
  webpage <- read_html(url)
  df = webpage %>% html_nodes("tr") %>% html_text()
  df = df[-c(1:5)]
  df = sub("\nMetro\n", "", df)
  df = sub("\n\n", "\n", df)
  data = do.call(rbind.data.frame, strsplit(df, "\n"))
  colnames(data) <- c("town", "Br0", "Br1", "Br2", "Br3", "Br4", "remove", "double")
  data = data %>% dplyr::select(-remove, -double)
  
  data$year = year
  if(dim(rentMA)[1] == 0) {
    rentMA <- data
  }
  else {
    rentMA <- rbind(rentMA, data)
  }
  
}

rentMA$Br0 = parse_number(rentMA$Br0) # parse number from string
rentMA$Br1 = parse_number(rentMA$Br1)
rentMA$Br2 = parse_number(rentMA$Br2)
rentMA$Br3 = parse_number(rentMA$Br3)
rentMA$Br4 = parse_number(rentMA$Br4)

rentMA = rentMA %>% mutate(town = stringr::str_replace(town, " city", "")) %>% mutate(town = str_trim(town, side="right"))

rentMA = rentMA %>% mutate(town = stringr::str_replace(town, "Barnstable Town", "Barnstable"))
rentMA = rentMA %>% mutate(town = stringr::str_replace(town, "Free Town", "Freetown"))
rentMA = rentMA %>% mutate(town = stringr::str_replace(town, "Province town", "Provincetown"))
rentMA = rentMA %>% mutate(town = stringr::str_replace(town, "Belcher town", "Belchertown"))


# url <- 'https://www.bostonmagazine.com/property/single-family-home-price-chart-2021/'

rentMA$town <- sub(" city|Town", "", rentMA$town, ignore.case = T)

rent_in_MA_2006_to_2022 = rentMA %>% tidyr::pivot_longer(cols = c(Br0, Br1, Br2, Br3, Br4), names_to = "house_type")

rent_in_MA_2006_to_2022$town = str_remove_all( rent_in_MA_2006_to_2022$town, " ")

write_rds(rent_in_MA_2006_to_2022, file = "data/rent_in_MA_2006_to_2022.rds")

DT::datatable(rent_in_MA_2006_to_2022, rownames = F)

```


<br/>

## Download population data

from https://donahue.umass.edu/documents/UMDI_V2021_MCD_Population_Estimates_Table_2022.07.02.xlsx


```{r population}

population = readxl::read_excel("data/UMDI_Appendix_A_MA_Subcounty_Pop_2010-2019.xlsx", sheet = 1, skip = 2 )
population = population %>% dplyr::select(1:14) %>% dplyr::select(-Census, -'Estimates Base')
colnames(population) <- c("town", "county", 2010:2019)
population.df =  population %>% tidyr::pivot_longer(!c("town", "county"), names_to = "year", values_to = "value")
population.df$year = as.integer(population.df$year)
colnames(population.df) <- c("town", "county", "year", "population")
write_rds(population.df, file = "data/populationMA.rds")
DT::datatable(population.df, rownames = F)

```
<br/>

## Teacher salary and school budget in MA school districts

Resource: https://profiles.doe.mass.edu/statereport/teachersalaries.aspx
The data is hard to scrape. It was copied/pasted from the resrouce above.



```{r teachers_salary}

salary = read_delim("data/teacher_salary_2019_2020.txt")
salary$total_budget = parse_number(salary$total_budget)
salary$average_salary = parse_number(salary$average_salary)
write_rds(salary, file = "data/teacher_salary_in_MA_2019.rds")
DT::datatable(salary, rownames = F)

```


<br/>

## Home owner ratio

Ratio of the home owner in the specific town is a potentially important factor for rental property. If this ratio is too high, maybe there is neither market or demand for rental seekers.
The data was scraped from http://archive.boston.com/yourtown/specials/snapshot/snapshot_percent_owner_occupied_housing/.

```{r}



url = 'http://archive.boston.com/yourtown/specials/snapshot/snapshot_percent_owner_occupied_housing/'
print(url)
#Reading the HTML code from the website
webpage <- read_html(url)
df = webpage %>% html_nodes("tr") %>% html_text()
df = df[-c(1:1)]
df = sub("\nMetro\n", "", df)
df = sub("\n\n", "\n", df)
data = do.call(rbind.data.frame, strsplit(df, "\n"))

colnames(data) <- "temp"

percent_of_homeowner = data %>% separate(temp, into = c("town", "num"), 
           sep = "(?<=[A-Za-z])(?=[0-9])") %>% separate(num, sep="%", into=c("value", "rank_based_on_homeowner"))

percent_of_homeowner$year = "2011"
percent_of_homeowner = percent_of_homeowner %>% dplyr::select(-rank_based_on_homeowner)
percent_of_homeowner$cat = "percent_of_homeowners"
percent_of_homeowner = percent_of_homeowner %>% dplyr::select(town, year, value, cat)
write_rds(percent_of_homeowner, file = "data/percent_of_homeowner.rds")

DT::datatable(percent_of_homeowner, rownames = F)

```

<br/>


## House prices in MA

Source: https://www.bostonmagazine.com/property/single-family-home-price-chart-2020/


```{r}


##########################################################
################ 2020 + 2016
##########################################################
url = "https://www.bostonmagazine.com/property/single-family-home-price-chart-2020/"
webpage <- read_html(url)
df = webpage %>% html_nodes("tr") %>% html_text()
df = df[-c(1:1)]
# df = sub("\n\n", "\n", df)
house_prices_ma = do.call(rbind.data.frame, strsplit(df, "\\$"))
colnames(house_prices_ma) <- c("town", "2020", "2019", "2018", "2015", "2014", "2010", "temp")
house_prices_ma = house_prices_ma  %>% dplyr::select(-temp, -`2010`)
# house_prices_ma %>% select(`2010`) %>% mutate(tem = str_sub(`2010`, start = 1, end = 7))
house_prices_slim = house_prices_ma %>% pivot_longer(!c("town"), names_to = "year", values_to = "value")
house_prices_slim$year = parse_integer(house_prices_slim$year)
house_prices_slim$value = parse_number(house_prices_slim$value)

##########################################################
################ 2021 + 2016 #############################
##########################################################
url = "https://www.bostonmagazine.com/property/single-family-home-price-chart-2021/"
webpage <- read_html(url)
df = webpage %>% html_nodes("tr") %>% html_text()
df = df[-c(1:1)]
house_prices_ma_2016_2021 = do.call(rbind.data.frame, strsplit(df, "\\$"))
colnames(house_prices_ma_2016_2021) <- c("town", "2021", "2020", "2016", "2011")
head(house_prices_ma_2016_2021)
house_prices_ma_2016_2021 = house_prices_ma_2016_2021 %>% dplyr::select(town, `2021`, `2016`)
house_prices_ma_2016_2021 = house_prices_ma_2016_2021 %>% pivot_longer(!c("town"), names_to = "year", values_to = "value")

house_prices_ma_2016_2021$year = parse_integer(house_prices_ma_2016_2021$year)
house_prices_ma_2016_2021$value = parse_number(house_prices_ma_2016_2021$value)

##########################################################
############### Merging #################################
##########################################################

house_prices_slim = rbind(house_prices_slim, house_prices_ma_2016_2021)
house_prices_slim = house_prices_slim %>% arrange(-year)
house_prices_slim$category = "single_family_home_price"

write_rds(house_prices_slim, file = "data/house_prices_MA.rds")

DT::datatable(house_prices_slim, rownames = F)



```

